---
title: "sunburst_motivation"
output: html_document
date: "2023-05-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('sunburst_utils.R')
source('sunburst_plot.R')
```

## Sunburst Motivation

Visualizing patterns in how people move across *states* can be a useful way to understand trends over time. In health sciences a researcher may be interested in how a patient with a disease gets treated for that disease, for example they may start on a treatment and transition to a different treatment if the initial treatment is not effective. Seeing these treatment patterns in aggregate can be a useful tool. 

To visualize this sort of information a starting point for what is needed of the data is a *subject*, a *state* (here, the treatment), and a *stage* (here the **time** component).
```{r echo=F}
df_long = data.frame(SUBJECT_ID = c(1,1,2,2,3,3,4,4,5,5),
                     TIME   = rep(1:2,5),
                     TREATMENT =  paste0('Drug ',c('A','B','B','A','A','A','A','A','B','C')))
df_long
```

Data of this form can the be aggregated at each stage. So in the above example 3 out of 5 people are on Drug A at Time 1. Of those 3 people, 1  transitions to Drug B at Time 2, while the other 2 does not transition to a different drug. 

```{r echo=F, message = F}
library(gt)
df_long %>%
  make_sequence(SUBJECT_ID,TREATMENT,TIME) %>%
  arrange(stage, stage_seq)  %>%
  select(Stage = stage, Sequence = stage_seq, Count = ct) %>%
  gt
```

A sunburst plot can allow you to see these pathways.

```{r echo=F, message = F}
plt <- df_long %>%
  make_sequence(SUBJECT_ID,TREATMENT,TIME) %>%
  prep_for_plot %>%
  ggsunburst_dynamic
girafe(ggobj = plt)

```

In this figure, the inner ring is 60% <span style="color: #F8766D;">coral</span> and 40% <span style="color: green;">green</span> to correspond with the proportions of subjects on those drugs during TIME = 1. Similarly each partition in the outer ring is either 20% or 40% to correspond to the percentage of patients that followed a given pathway.


